---
title: "BSMM-lab-1"
subtitle: "BSMM 8740 Fall 2023"
author: "Harsha"
date: "22-09-2023"
format: html
editor: visual
self-contained: true
---

## Setup

Load packages and data:

```{r load-pkg-data}
#| message: false
library(Lahman)
```

## Exercises

### Exercise 1

```{r glimpse-data}
dplyr::glimpse(Teams)
```

The `Teams` dataset has 3015 observations and 48 variables.

### Exercise 2

```{r}
mets <- Teams  |>
  dplyr::filter(teamID == "NYN")
my_mets <- mets |>
  dplyr::filter(yearID %in% 2004:2012) |>
  dplyr::select(yearID,G,W,L)
my_mets 
```

### Exercise 3

...
```{r}
mets_ben <- Teams |>
  dplyr::select(yearID,teamID,G,RA,RS = R) |>
  dplyr::filter(teamID == "NYN" & yearID %in% 2004:2012)
mets_ben
```

### Exercise 4

```{r}
mets_ben <- mets_ben |>
   dplyr::mutate(WPct = W / (W + L))
mets_ben
```

```{r}
mets_ben <- mets_ben |>
   dplyr::mutate(W_hat = WPct_hat * (W + L))
mets_ben
...

### Exercise 5

```{r}
filter(mets_ben, W >= W_hat)
...

```{r}
filter(mets_ben, W < W_hat)
...

### Exercise 6

...
```{r}
dplyr::arrange(mets_ben, desc(WPct))
...

### Exercise 7

...
```{r}
mets_ben %>% 
  dplyr::mutate(Diff = W - W_hat) |>
  dplyr::arrange(desc(Diff))
...

### Exercise 8

...
```{r}
mets_ben |>
  dplyr::summarize(
    num_years = n(), 
    total_W = sum(W), 
    total_L = sum(L), 
    total_WPct = sum(W) / sum(W + L), 
    sum_resid = sum(W - W_hat)
  )
...

In these nine years, the Mets had a combined record of 728 wins and 730 losses, for an overall winning percentage of 0.499.

### Exercise 9 \_ Data not Available

### Exercise 10

```{r}
library(dplyr)
mets_ben <- mets_ben %>%
  dplyr::mutate(
    gm = case_when(
      yearID == 2004 ~ "Jim Duquette",
      yearID >= 2005 & yearID <= 2010 ~ "Omar Minaya",
      yearID >= 2011 ~ "Sandy Alderson"
    )
  )
mets_ben
```

##Exercise 11

```{r Group by the manager (gm) variable and summarize the data}
 
manager_summary <- mets_ben %>%
  group_by(gm) %>%
  summarize(
    num_years = n(),
    total_W = sum(W),
    total_L = sum(L),
    total_WPct = mean(WPct),
    sum_resid = sum(Diff)
  )
```

```{r}
manager_summary
```
